A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-18 text/introduction.tex
FInish revising Seciton 1.

Multi-stage programming enables us to generate and run fragments of code at runtime.  \AI{It sounds like the subject of ``generate and run'' is human being (``us'').  Something like ``Multi-stage programming offers linguistic support for dynamic code generation and execution.''?}
MetaOCaml~\cite{oleg2014} \AI{You need a space between a word and a citation.  Taha's original MetaOCaml should be cited, too.} provides features for multi-stage programming, which are brackets, escape, and run.  \AI{CSP?}
Brackets, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Escape, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, escape is supposed to use in a fragment of code, that is, escape cannot be used as an alternative of run.
For example, the following MetaOCaml expression

@@

% \subsubsection{ベクトル計算も高速化できる}

Multi-stage programming can also optimize vector calculation.
For example, \verb|vadd| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|vadd| function for a given vector length and optimize it with multi-stage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

% \subsubsection{生成した行列計算のコードは特定のサイズに特化しており、他のサイズでは使えない。
% このため、間違って組み合わせると容易にエラーを引き起こす。既存の型システムではこのエラーを防止できない}

However, there is a sever problem in functions which are optimized by multi-stage programming.
Although unrolled \verb|vadd| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
\AI{Why SEGV?  MetaOCaml will raise an exception.}
This problem is serious but existing type systems for multi-stage calculi cannot prevent it.

% We call this problem \red{Specialized Code Problem}.

\AI{このイントロ、各段落はまあまあよいのですが、全体の流れが、ふたつ目の文から MetaOCaml という特定の多段階プログラミング言語の話をして、それから応用と問題点の話になりますが、ここが多段階プログラミング一般の話なのか MetaOCaml 固有の話なのかがぶれているように思います。まずは、多段階プログラミング一般の話を応用も含めてする。(他の多段階計算ができるシステムを無視しない。)それから、MetaOCaml の紹介と、その型システムの問題点の話をすべきでしょう。MetaOCaml の型システムでできることもふれておいた方がよいでしょう。}

@@

As we pointed out in the above section,
functions optimized with multi-stage programming can take only restricted values as arguments.
When you optimize \verb|vadd| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multi-stage calculus
so that the type system can guarantee optimized functions are used properly.

In this paper, we design a new multi-stage calculus \LMD by 
merging dependent types into existing multi-stage calculus \LTP\cite{Hanada2014}.
\LMD is multi-stage calculus which contains dependent types.
\LMD makes multi-stage programming safer with the power of dependent types.
Our technical contribution are summarized as follows:
\begin{itemize}
        \item we give the formal definition of \LMD with its syntax, 
        type system which contains type equivalence, full, reduction, and staged reduction;
        \item for the full reduction, we prove subject reduction, strong normalization, and confluence
        \item fot the staged reduction, we prove unique decomposition and progress.
\end{itemize}

@@

The organization of this paper is the following.
Section 2 gives an informal overview of \LMD.
Section 3 defines \LMD and
Section 4 shows the properties of \LMD including Unique Decomposition \AI{There are more important properties that should be mentioned!  In fact, the properties should be mentioned earlier.}.
Section 5 discusses future works.
Finally, Section 6 discusses related works.
\AI{Use $\backslash$label and $\backslash$ref!}
\AI{``Related work'' and ``future work'' (they are uncountable).}
\AI{Related work is discussed in Section 5?  Section 6 is the section for conclusion, which includes future work.}

>>>>

Multi-stage programming enables programmers to generate and run fragments of code at runtime.
We can treat a piece of code as a value with multi-stage programming.
It means code fragments can be substituted into variables, combined and run at runtime.
Multi-stage programming improves the performance of programs by generating optimized code for a given input at runtime. \cite{taha2007gentle}
% MetaOCamlを例に挙げて多段階計算について説明する

MetaOCaml~\cite{calcagno2003implementing}\cite{oleg2014} provides features for multi-stage programming, which are brackets, escape, run, and CSP.
Brackets, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Escape, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, escape is supposed to use in a fragment of code, that is, escape cannot be used as an alternative of run.
For example, the following MetaOCaml expression

@@

% 多段階計算の型理論の既存研究の紹介

Existing type systems for multi-stage calculi ensure that
the generated code is well-typed and code generation and evaluation are performed in the intended order.
For example, the type system of MetaOCaml~\cite{taha1999multi} classify expressions in a program
with the depth of brackets in order to handle them appropriately.
In $\lambda^\triangleright$ by Tsukada and Igarashi~\cite{Tsukada},
they use environment classifier to classify and 
succeeded in recognizing code values which are run safely at type-checking time.

% ベクトル計算も高速化できる

Multi-stage programming can also optimize vector calculation.
For example, \verb|vadd| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|vadd| function for a given vector length and optimize it with multi-stage programming.

% 多段階計算で生成したコードは特殊化されているが故に問題も多い

However, there is a sever problem in functions which are optimized by multi-stage programming.
Although unrolled \verb|vadd| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a exception.
This problem is serious but existing type systems for multi-stage calculi cannot prevent it.

@@

As we pointed out in the above section,
functions optimized with multi-stage programming can take only restricted values as arguments.
When you optimize \verb|vadd| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multi-stage calculus
so that the type system can guarantee optimized functions are used properly.
Although there is a trial to combine multi-stage programming and dependent type by Brady and Hammond~\cite{brady2006dependently},
they didn't give a formal definition and properties for the calculus.

In this paper, we design a new multi-stage calculus \LMD by 
merging dependent types into existing multi-stage calculus \LTP\cite{Hanada2014}.
\LMD is multi-stage calculus which contains dependent types.
\LMD makes multi-stage programming safer with the power of dependent types.
Our technical contribution are summarized as follows:
\begin{itemize}
        \item we give the formal definition of \LMD with its syntax, 
        type system which contains type equivalence, full, reduction, and staged reduction;
        \item for the full reduction, we prove preservation, strong normalization, and confluence
        \item fot the staged reduction, we prove unique decomposition and progress.
\end{itemize}

@@

The organization of this paper is the following.
Section \ref{sec:informal-overview} gives an informal overview of \LMD.
Section \ref{sec:formal-definition} defines \LMD and
Section \ref{sec:properties} shows the properties of \LMD including preservation, strong normalization, confluence, and progress.
Section \ref{sec:related-work} discusses related work.
Finally, we conclude this paper in Section \ref{sec:conclusion} and discusses future work.
